---
title: "Forecasting covid-19 cases by state using Time-Series"
output:
  html_document: 
    theme: cerulean
---
## We use WV as a test state for tunning time series models.
```{r setup, include=FALSE}

library("forecast")
library("tseries") 	
library("prophet")

```

### Import data
```{r }

covid <- read.csv("United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")
colnames(covid) <- c("date","KS", "WV")

```

plot and data cleansing
```{r}
wv = covid$WV
wv <- na.omit(wv)
n <- length(wv)
ts.wv <- ts(wv, frequency = n, end = c(2022,2,9))
plot.ts(ts.wv)
```

### Auto Arima -- Not giving good MAPE

- The curve does not go smoothly, so we will try extract only the last 'smooth' part later
```{r}
auto <- auto.arima(ts.wv)		# fits ARIMA(p,d,q) x (P, D, Q) automatically
summary(auto)
auto.predict <- forecast:::forecast.Arima(auto, h = 8, level = c(68, 90))
plot(auto.predict)
```



### ARIMA by hand
adf.test for differences = 2 < 0.1, so it suggests differentiate twice
```{r}
adf.test(ts.wv)
wv.d1 <- diff(ts.wv, diffrences = 1)
adf.test(wv.d1)

wv.d2 <- diff(ts.wv, differences = 2)
adf.test(wv.d2)
wv.d2 <- na.omit(wv.d2)
```
Determine p, q by looking at plots.  
The values converges very slow.  
```{r}
# AR(p) => Pacf 
Pacf(wv.d2, lag.max = 50)	
# MA(q) => Acf
Acf(wv.d2, lag.max = 30)	
```

Use different p,q and d = 2 to build ARIMA model, compare MAPE, the best one is ARIMA(6,2,3) with MAPE of 0.2537424  
```{r}
library(forecast)
h = 7
# far2 <- function(x, h){forecast(Arima(ts.wv, order = c(6,2,2), method="ML"), h=h)}
# e <- tsCV(ts.wv, far2, h=7)
#res.m1 <- summary(m1)
m1 <- Arima(ts.wv, order = c(6,2,2), method="ML")
res.m1 <- summary(m1)
res.m1

m2 <- Arima(ts.wv, order = c(6,2,3), method="ML")
res.m2 <- summary(m2)
res.m2

m3 <- Arima(ts.wv, order = c(7,2,3), method="ML")
res.m3 <- summary(m3)
res.m3
```
### Prophet
```{r}
library(prophet)
library(lubridate)

## prepare data to meet prophet model standard
wv.p <-  covid[,c("date", "WV")]
wv.p <- na.omit(wv.p)
colnames(wv.p) <- c("ds","y")
# convert col to 'date' type
wv.p$ds <- mdy(wv.p$ds)
m6 <- prophet(wv.p)

## make predictions
m6.predict <- make_future_dataframe(m6, periods = 7)
forecast <- predict(m6, m6.predict)
tail(forecast[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')], n=7)
df.cv <- cross_validation(m6, horizon = 7, units = 'days')
df.p <- performance_metrics(df.cv)
head(df.p)

```


As said at the beginning, we tried with data after the last disrupt change in slope, which is the early Jan of 2022, with the guess that omicron occurred.   

## Build model with date > 1/3/2022  
```{r}
covid.2 <- tail(covid,n=38)
```

Running tests for ARIMA, and we can see the line is more smooth, compared to the one above:  
```{r}
wv.2 = covid.2$WV
wv.2 <- na.omit(wv.2)
n.2 <- length(wv.2)
ts.wv.2 <- ts(wv.2, frequency = n.2, end = c(2022,2,9))
plot.ts(ts.wv.2)
adf.test(ts.wv.2)
wv.2.d1 <- diff(ts.wv.2, diffrences = 1)
adf.test(wv.2.d1)

wv.2.d2 <- diff(ts.wv.2, differences = 2)
adf.test(wv.2.d2)

Pacf(wv.2.d2, lag.max = 10)	
# MA(q) => Acf
Acf(wv.2.d2, lag.max = 10)	
```

Using ARIMA, setting p,d,q by hand:
```{r}
h = 7
# far2 <- function(x, h){forecast(Arima(ts.wv, order = c(6,2,2), method="ML"), h=h)}
# e <- tsCV(ts.wv, far2, h=7)
#res.m1 <- summary(m1)
m4 <- Arima(ts.wv.2, order = c(4,2,4), method="ML")
res.m4 <- summary(m4)
res.m4

######### best ARIMA model
m5 <- Arima(ts.wv.2, order = c(4,2,7), method="ML")
res.m5 <- summary(m5)
res.m5

```

Using Prophet  
**This model has the least MAPE of all models!** We finally adopted this.
```{r}
wv.p.2 <-  covid.2[,c("date", "WV")]
wv.p.2 <- na.omit(wv.p.2)
colnames(wv.p.2) <- c("ds","y")
# convert col to 'date' type
wv.p.2$ds <- mdy(wv.p.2$ds)
m7 <- prophet(wv.p.2)

## make predictions
m7.predict <- make_future_dataframe(m7, periods = 10)
forecast.m7 <- predict(m7, m7.predict)
tail(forecast.m7[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')], n=7)
df.cv.m7 <- cross_validation(m7, horizon = 7, units = 'days')
df.p.m7 <- performance_metrics(df.cv.m7)
head(df.p.m7)
```